---
title: "Chapters 4 and 5 Homework"
author: "Your Name Here"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document:
    reference_docx: HWSTY.docx
    highlight: tango
  pdf_document: default
---

```{r setup, include=FALSE}
if(!require(pacman)) install.packages("pacman")
library("pacman")
pacman::p_load("tidyverse","psych","jtools","car","Rmisc","knitr","rmarkdown")

knitr::opts_chunk$set(echo = TRUE, render = knitr::normal_print, 
                      comment=NA, attr.source='.numberLines')
```

# Problem 1.

*This problem was adapted from: Chapter 4 problem 4.28 (Principles of econometrics. John Wiley & Sons; 2018).*

The file `data/wa_wheat.csv` contains observations on wheat yield in Western Australian shires. There are 48 annual observations for the years 1950-1997. For the Northampton shire, consider the following four equations:

-   mod1: $YIELD_t = \beta_0+\beta_1 TIME + e_t$
-   mod2: $YIELD_t = \beta_0+\beta_1 \log(TIME) + e_t$
-   mod3: $YIELD_t = \beta_0+\beta_1 TIME^2 + e_t$
-   mod4: $\log(YIELD_t) = \beta_0+\beta_1 TIME + e_t$

### A. Estimate each model.

#### Load and inspect the data

1.  Use `read.csv()` to load the `data/wa_wheat.csv` into a data frame named wheat.
2.  View `data/wa_wheat.def` by clicking on the file in the Files tab.
3.  Use `head()` to inspect the first few rows of the data.
4.  Use `psych::describe()` to display a table of summary statistics.

```{r}

```

#### Estimate models one through four and display summaries of each.

1.  Use `lm()` to estimate each model. Note that the yield for Northampton shire is stored in the series `northampton`.
2.  Use `jtools::export_summs()` to display a summary of each model in a single table.
3.  Use the `model.names` argument to name models 1 through 4 "Linear", "Linear-Log", "Quadratic", "Log-Linear", respectively.

```{r}


```

### B. Draw plots of actual and fitted values vs. time, and the residuals vs. fit plots for each model.

#### I have provided you with two functions, `plot.time()` and `plot.fitted()`, to help draw the required charts.

```{r plot_fitted, echo=F}

# These two functions will help you. plot.time() will draw the actual and fitted 
# values vs. time, and plot.fitted() will draw the residuals vs. fit plot. Both 
# functions need only a linear model as an argument.

plot.time <- function (lmod, ln=F) {
  require(ggplot2,dplyr)
  pr <- data.frame(
    Observation = seq(from = 1, to = nrow(lmod$model)),
    Actual = if(ln==T) exp(lmod$model[, 1]) else lmod$model[, 1],
    Fitted = if(ln==T) exp(fitted(lmod)+(summary(lmod)$sigma^2)/2) else fitted(lmod)
  ) %>% pivot_longer(.,
                     cols = 2:ncol(.),
                     names_to = "series",
                     values_to = "values")  %>%
    ggplot(aes(x = Observation, y = values, color = series)) +
    geom_line() +
    ggtitle(lmod$call) +
    theme_classic() +
    scale_color_manual(values = c("#2c7bb6", "#d7191c", "#808080", "#000000")) +
    theme(
      legend.position = "bottom",
      legend.direction = 'horizontal',
      legend.title = element_blank()
    )
  return(pr)
}

plot.fitted <- function(lmod) {
  require(ggplot2,dplyr)
  df = lmod$model
  title = paste("Residuals vs. Fits (Dependent Variable: ",
                names(df)[1],
                ")",
                sep = "")
  pr <- df %>% ggplot(aes(x = fitted(lmod), y = resid(lmod))) +
    geom_ribbon(aes(ymin = -sigma(lmod), ymax = sigma(lmod)),
                fill = "gray",
                alpha = .5) +
    geom_ribbon(aes(ymin = -sigma(lmod) * 2, ymax = sigma(lmod) * 2),
                fill = "lightgray",
                alpha = .5) +
    geom_point() +
    theme_classic() +
    ggtitle(title) +
    ylab("Residules") +
    xlab("Fitted") +
    geom_hline(yintercept = 0,
               linetype = "solid",
               color = "black")
  return(pr)  
}  
```

#### Use `Rmisc::multiplot()` and `plot.time()` to draw plots of actual and fitted values vs. time. Use `ggtitle()` to label the plot for each model "Linear", "Linear-Log", "Quadratic", and "Log-Linear" e.g. `plot.time(mod1) + ggtitle("Linear")`.

```{r}


```

#### Use `Rmisc::multiplot()` and `plot.fitted()` to draw plots of residual vs. fitted values. Label each plot as you did before.

```{r echo=T}


```

### C. Calculate the generalized $R^2$ for each model.

#### Use the `cor()` to calculate the generalized $R^2$. Use `data.frame()` to store the data values in a data frame, labeling each the same as the model title.

```{r echo=T}


```

### D. Based on your analysis, which of these models is the best. Why?

[Your answer here]

# Problem 2.

*This problem was adapted from: Chapter 5 problems 5.21 and 5.22 (Principles of Econometrics. John Wiley & Sons; 2018).*

Using the data in the file `data/toody5.csv`, estimate the model

$${Y_t} = {\beta _1} + {\beta _2}TREND_t + {\beta _3}RAIN_t + {\beta _4}RAIN_t^2 + {\beta _5}\left( {RAIN_t \times TREND_t} \right) + {e_t}$$ where $Y_t$ = wheat yield in tons per hectare in the Toodyay Shire of Western Australia in year $t$; $TREND_t$ is a trend variable designed to capture technological change; $RAIN_t$ is total rainfall in decimeters (dm) from May to October (the growing season) in year $t$.

### A. Estimate the model.

#### Load the data into a data frame named `toodyay`.

```{r echo=T}


```

#### Use `lm()` to estimate the model, store the model in an object named `toodyay.lm`, and use `jtools:summ()` to present a summary of the model.

```{r echo=T}


```

### B. Evaluate the model.

#### Plot actual and fitted values vs. time. Use `ggtitle()` to lable the plot "Wheat yield for the Toodyay Shire of Western Australia."

```{r}


```

#### Evaluate the overall fit

[Your answer here]

#### Plot the residual vs. fitted values.

```{r}


```

#### Evaluate residual vs. fitted values plot.

[Your answer here]

## C. Estimate a log-linear form of this model, i.e.

$$\log {Y_t} = {\beta _1} + {\beta _2}TREND_t + {\beta _3}RAIN_t + {\beta _4}RAIN_t^2 + {\beta _5}\left( {RAIN_t \times TREND_t} \right) + {e_t}$$

#### Estimate and summarize the model.

```{r echo=T}


```

#### Plot actual and fitted values vs. time. Remember to label the plot.

```{r}


```

#### Plot the residual vs. fitted values.

```{r}


```

#### Calculate the generalized $R^2$ for both the linear and log models, and store those in a data frame.

```{r echo=T}


```

### D. Evaluate which model is the best.

[Your answer here]
